Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-04-29 code-tests.tex
oops, needed to do a bit of cleanup in 1-zone test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{One-zone collapse test}
\label{sec.tests.1-zone}
The one-zone collapse test simulates the collapse of a
self-gravitating gas cloud using a semi-analytic model for the
evolution of gas density and adiabatic heat input as a function of
time.  It is designed to test the chemistry and cooling modules over a
wide range in densities and over physically motivated timescales.
Because this test disables the hydrodynamic and gravity solvers and
uses a simple model for the density evolution, it is far faster than
running a true collapse simulation.  The density evolution is based on
the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
  1969MNRAS.144..425P} solution for isothermal collapse with a
modification to account for the efficiency with which the heat
introduced by compression can be radiated away
\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
here, follows the work of \citet{2005ApJ...626..627O}, and we direct
the interested reader to this paper for further details.  The
evolution of the gas density, $\rho$, is given by 
\begin{equation}
\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
\end{equation}
where the collapse timescale, $t_{col}$, is
\begin{equation} \label{eqn.tcol}
t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
\end{equation}
In this equation, $t_{dyn}$ is the dynamical time and is expressed as
\begin{equation}
t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
\end{equation}

with G being the gravitational constant.  The collapse timescale is altered from the dynamical time by a
factor $f$ in Equation \ref{eqn.tcol}, which is an approximation of the
ratio of the gas pressure to the force of gravity.  The value of $f$
depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
ln\ p / \partial ln\ \rho)$, which we calculate by linearly extrapolating
from derivative values for the two previous timesteps.  For the value
of $f$ in this test problem, we use the piecewise function of \citet{2005ApJ...626..627O},
given by
\begin{equation}
f = \left\{
  \begin{array}{ll}
  0, & \gamma_{ef} < 0.83,\\
  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
  \gamma_{ef} < 1,\\
  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
\end{array} \right.
\end{equation}
The specific energy evolves as
\begin{equation}
\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
\end{equation}
where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
and energy, temperature, density, and pressure are related by the
ideal gas law.  Figure \ref{fig.onezone} shows an example of the
one-zone collapse test performed with an initial number density of 1
hydrogen atom per
cm$^{-3}$ and an initial temperature of 100 K using the 12 species chemistry
network with H, D, and He species and metal cooling rates calculated
with the \texttt{Cloudy} code.  The effects of metal cooling can be
clearly seen; as the metallicity increases from zero to
$10^{-2}$~Z$_\odot$, the gas deviates from the primoridal result
(black line) both more rapidly and more significantly.  Our primordial results
compare very well to those shown in \citet{2005ApJ...626..627O}; however,
we use a different cooling method than Omukai, so the lines describing
the evolution of the metal-enriched gas are not directly comparable.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{One-zone collapse test}
\label{sec.tests.1-zone}

The one-zone collapse test simulates the collapse of a
self-gravitating gas cloud using a semi-analytic model for the
evolution of gas density and adiabatic heat input as a function of
time.  It is designed to test the chemistry and cooling modules over a
wide range in densities and over physically motivated timescales.
Because this test disables the hydrodynamic and gravity solvers and
uses a simple model for the density evolution, it is far faster than
running a true collapse simulation.  The density evolution is based on
the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
1969MNRAS.144..425P} solution for isothermal collapse with a
modification to account for the efficiency with which the heat
introduced by compression can be radiated away
\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
here, follows the work of \citet{2005ApJ...626..627O}, and we direct
the interested reader to this paper for further details.  The
evolution of the gas density, $\rho$, is given by

\begin{equation}
\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
\end{equation}

where the collapse timescale, $t_{col}$, is

\begin{equation} \label{eqn.tcol}
t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
\end{equation}
In this equation, $t_{dyn}$ is the dynamical time and is expressed as
\begin{equation}
t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
\end{equation}

with G being the gravitational constant.  The collapse timescale is
altered from the dynamical time by a factor $f$ in Equation
\ref{eqn.tcol}, which is an approximation of the ratio of the gas
pressure to the force of gravity.  The value of $f$ depends on the
effective adiabatic index, $\gamma_{ef} \equiv (\partial ln\ p
/ \partial ln\ \rho)$, which we calculate by linearly extrapolating
from derivative values for the two previous timesteps.  For the value
of $f$ in this test problem, we use the piecewise function of
\citet{2005ApJ...626..627O}, given by

\begin{equation}
f = \left\{
  \begin{array}{ll}
  0, & \gamma_{ef} < 0.83,\\
  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
  \gamma_{ef} < 1,\\
  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
\end{array} \right.
\end{equation}

The specific energy evolves as

\begin{equation}
\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
\end{equation}

where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
and energy, temperature, density, and pressure are related by the
ideal gas law.  Figure \ref{fig.onezone} shows an example of the
one-zone collapse test performed with an initial number density of 1
hydrogen atom per cm$^{-3}$ and an initial temperature of 100 K using
the 12 species chemistry network with H, D, and He species and metal
cooling rates calculated with the \texttt{Cloudy} code.  The effects
of metal cooling can be clearly seen; as the metallicity increases
from zero to $10^{-2}$~Z$_\odot$, the gas deviates from the primoridal
result (black line) both more rapidly and more significantly.  Our
primordial results compare very well to those shown in
\citet{2005ApJ...626..627O}; however, we use a different cooling
method than Omukai, so the lines describing the evolution of the
metal-enriched gas are not directly comparable.
